---
title: "Lab 6 Comparing two means"
author: "Jonathan Kanyinda"
date: "2021-04-22"
output: github_document
---


Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

## Question A

> What is the mean different in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this mean difference. Show your code and write a sentence giving your answer

ANSWER

```{r}

# load packages -----------------------------------------------------------

library(tidyverse)


# read data ---------------------------------------------------------------

fish <- read_csv("chap12q19ElectricFish.csv")

# put data in tidy format ------------------------------------------------

fish_long <- 
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>% 
  mutate(location = str_remove(location, c("species"))) %>% 
  print()

t.test(formula = species ~ location, data = fish_long)
```

 Mean difference :

Mean_difference = 16.41667 - 14.58333

Mean_difference # The mean difference is 1.83334

 95% confidence intervall of Mean length difference :

Upper_limit = -4.587031

Lower_limit = 8.253697


## Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

ANSWER

```{r}
t.test(formula = species ~ location, data = fish_long)
```

The p value 0.5596 is greater than  0.05, therefore the null hypothesis will not be rejected. This is because the null hypotheis is rejected when the p value is smaller than 0.05, and since the obtained p value was 0.5596 which was greater than 0.05, there was failure to reject the null hypothesis stating that there was no difference in teh number of species in the main rivers caused by the presence of tributaries.
The means from the different locations were different, the mean in the downstream group was  16.41667 and the mean in the upstream group was 14.58333. The mean difference was 1.83334

## Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

ANSWER

```{r}
fish_long %>% 
  ggplot(aes(x = species)) +
  geom_histogram(
    aes(fill = location), 
    bins = 15, 
    alpha = 0.5, 
    position = "identity"
  ) +
  scale_fill_manual(values = c("darkorange","cyan4")) +
  theme_minimal()
```




For Question A, the assumption made was that the presence of tributaries did not affect the local
number of electric fish in the main rivers (Also the Null hypothesis). For part B, the assumption 
was that the data was normally distributed and that the peaks/umps of the two locations 
would be close enough to generate an overall normal distribution.
Based on the graph,the distribution does not appear to be normal and therefore
does not meet teh assumption made about the normality.
